---
title: "finc584_ps6"
author: "William Lee"
output: pdf_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)

xfun::pkg_attach2('tidyverse', 'readxl', 'sandwich', 'lmtest', 'ivreg', 'knitr')

proj <- dirname(rstudioapi::getSourceEditorContext()$path)
```

## 1 A

The naive point estimates are presented in part B for simplicity.  

```{r 1a}

fred <- readxl::read_xlsx(file.path(proj, "FRED-QD.xlsx")) %>% 
  select(pcecc96, sp500) %>% 
  mutate(R_t = sp500/dplyr::lag(sp500,1), 
         R_t1 = dplyr::lag(R_t,1),
         R_t2 = dplyr::lag(R_t,2),
         c_ratio = pcecc96/dplyr::lag(pcecc96,1),
         c_ratio1 = dplyr::lag(c_ratio,1),
         c_ratio2 = dplyr::lag(c_ratio,2)) %>% 
  drop_na()

gmm.moments <- function(params, data = fred){
  beta <- params[1]
  gamma <- params[2]
  
  moment1 <- beta*(data['c_ratio']^(-gamma))*data['R_t'] - 1
  moment2 <- moment1 * data['R_t1']
  moment3 <- moment1 * data['R_t2']
  moment4 <- moment1 * data['c_ratio1']
  moment5 <- moment1 * data['c_ratio2']
  
  return(as.matrix(cbind(moment1, moment2, moment3, moment4, moment5)))
}

l2_metric <- function(x, weight_mat){
  return(colMeans(gmm.moments(x)) %*% weight_mat %*% colMeans(gmm.moments(x)))
}

naive <- optim(par = c(.7,2), fn = function(params) 
  l2_metric(params, weight_mat = diag(5)), lower = c(0,0), upper = c(1, 20))


```
## 1 B


```{r 1b}

w_optimal <- solve(1/nrow(fred) * t(gmm.moments(naive$par)) %*% gmm.moments(naive$par))

optimal_params <- optim(par = naive$par, fn = function(params) 
  l2_metric(params, weight_mat = w_optimal), lower = c(0,0), upper = c(1, 20))

kable(data.frame(var = c('beta', 'gamma'), 
                 naive = naive$par, optimal = optimal_params$par), digits = 4)

```

## 1 C




```{r 1c}

gradient <- function(params,  data = fred){
  beta <- params[1]
  gamma <- params[2]
  
  # Derivative WRT beta
  e <- data[['R_t']]*data[['c_ratio']]^(-gamma) 
  moment1_beta <- mean(e)
  moment2_beta <- mean(e * data[['R_t1']])
  moment3_beta <- mean(e * data[['R_t2']])
  moment4_beta <- mean(e * data[['c_ratio1']])
  moment5_beta <- mean(e * data[['c_ratio2']])
  
  # Derivative WRT gamma
  e <- -beta*data[['R_t']]*(data[['c_ratio']]^(-gamma))*log(data[['c_ratio']]) 
  moment_1_gamma <- mean(e)
  moment2_gamma <- mean(e * data[['R_t1']])
  moment3_gamma <- mean(e * data[['R_t2']])
  moment4_gamma <- mean(e * data[['c_ratio1']])
  moment5_gamma <- mean(e * data[['c_ratio2']])
  
  return(matrix(c(moment1_beta, moment2_beta, moment3_beta,
                  moment4_beta, moment5_beta, moment_1_gamma, moment2_gamma,
                  moment3_gamma, moment4_gamma, moment5_gamma), ncol = 5, byrow = T))

}

H <- gradient(optimal_params$par)

vcov_HC <- sqrt(diag((1/nrow(fred))* solve(H %*% w_optimal %*% t(H))))

ci.upper <- optimal_params$par + 1.96*(vcov_HC)
ci.lower <- optimal_params$par - 1.96*(vcov_HC)

kable(x = data.frame(var = c('beta', 'gamma'), optimal = optimal_params$par, 
                     se = vcov_HC, ci.lower = ci.lower, ci.upper = ci.upper), 
      digits = 3)


```


## 1 D

```{r 1d}

J <- nrow(fred) * colMeans(gmm.moments(optimal_params$par), w_optimal) %*% 
  w_optimal %*% colMeans(gmm.moments(optimal_params$par), w_optimal) 

J.Stat <- 1 - pchisq(J, df = 3)

```

The test statistic for over-identification is `r J` which comes from a $\chi^2(3)$. The corresponding p-value is `r J.Stat` which allows us to reject the null. The fact that instruments, which we think should have worked, are not satisfying the overidentification test make us question the validity of the model itself. 



